State of the Practice for Medical Imaging Software

### 2022-03-24 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Edits completed for Section 5 on Comparison to Community Ranking

To address \rqref{RQ_CompareHQ2Popular}, Table~\ref{tab_ranking_vs_GitHub}
compares our LBM software package rankings against their popularity in the
research community, as estimated by repository stars and watches. We answer this
question by grading the qualities of the software, collecting ratings from the
MI community, then conducting two comparisons:
\begin{itemize}
\item comparing our ranking with the community ratings on GitHub, such as GitHub
stars, number of forks, and number of people watching the projects (Section~\ref{sec_grading_software});
\item comparing top-rated software designated by our methodology with the ones
recommended by our domain experts (Section \ref{sec_software_selection}).
\end{itemize}

Table~\ref{tab_ranking_vs_GitHub} shows our ranking of the 29 MI projects, and
their GitHub metrics, if applicable. As mentioned in
Section~\ref{sec_score_maintainability}, 24 projects used GitHub. Since GitHub
repositories have different creation dates, we collect the number of months each
stayed on GitHub, and calculate the average number of new stars, people
watching, and forks per 12 months. The method of getting the creation date is
described in Section~\ref{sec_grading_software}.  We obtained these
measures in July, 2021.

In Table~\ref{tab_ranking_vs_GitHub}, we used the average number of new stars
per 12 months as an indicator of the GitHub popularity, and listed the items in
the descending order of this number. We ordered the non-GitHub items by our
ranking. Generally speaking, most of the top-ranking MI software projects also
received greater attention and popularity on GitHub. Between our ranking and the
GitHub stars-per-year ranking, four of the top five software projects are the
same.

Project \textit{dwv} was popular on GitHub, but we ranked it low. As mentioned
in Section~\ref{sec_result_installability}, we failed to build it locally, and
used the test version on its websites for the measurements. We followed the
instructions and tried to run the command ``yarn run test'' locally, which did
not work. In addition, the test version did not detect a broken DICOM file and
displayed a blank image as described in Section \ref{sec_result_robustness}. We
might underestimate the scores for \textit{dwv} due to uncommon technical
issues. We also ranked \textit{DICOM Viewer} much lower than its popularity. As
mentioned in Section \ref{sec_result_installability} it depended on the
NextCloud platform that we could not successfully install. Thus, we might
underestimate the scores of its surface reliability and surface
robustness. We weighted all qualities equally, which is not likely to be the
case with all users. As a result, some projects with high popularity may not
perform well in all qualities.

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
Software & Our ranking & Stars/yr & Watches/yr & Forks/yr \\ 
\midrule
3D Slicer & 1 & 284 & 19 & 128 \\
OHIF Viewer & 3 & 277 & 19 & 224 \\
dwv & 23 & 124 & 12 & 51 \\
ImageJ & 2 & 84 & 9 & 30 \\
ParaView & 5 & 67 & 7 & 28 \\
Horos & 12 & 49 & 9 & 18 \\
Papaya & 17 & 45 & 5 & 20 \\
Fiji & 4 & 44 & 5 & 21 \\
DICOM Viewer & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 40 & 4 & 17 \\
Weasis & 6 & 36 & 5 & 19 \\
dicompyler & 26 & 35 & 5 & 14 \\
OsiriX Lite & 9 & 34 & 9 & 24 \\
MRIcroGL & 18 & 24 & 3 & 3 \\
GATE & 25 & 19 & 6 & 26 \\
Ginkgo CADx & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 8 & 18 & 5 & 7 \\
Drishti & 27 & 16 & 4 & 4 \\
Slice:Drop & 20 & 10 & 2 & 5 \\
ITK-SNAP & 15 & 9 & 1 & 4 \\
medInria & 7 & 7 & 3 & 6 \\
SMILI & 13 & 3 & 1 & 2 \\
MatrixUser & 28 & 2 & 0 & 0 \\
MicroView & 16 & 1 & 1 & 1 \\
Gwyddion & 11 & n/a & n/a & n/a \\
XMedCon & 19 & n/a & n/a & n/a \\
DicomBrower & 21 & n/a & n/a & n/a \\
AMIDE & 22 & n/a & n/a & n/a \\
3DimViewer & 24 & n/a & n/a & n/a \\ 
\bottomrule
\end{tabular}
\caption{\label{tab_ranking_vs_GitHub}Software ranking versus GitHub metrics
(Sorted in descending order of GitHub popularity, as estimated by the number of
new stars per year)}
\end{table}
\endgroup

As shown in Section~\ref{sec_software_selection}, our domain experts recommended
a list of top software with 12 software products.  All of the top 4 software
from the Domain Expert's list are among the top 12 ranked by our methodology. 3
of the top 4 on both lists are the same: \textit{3D Slicer}, \textit{ImageJ},
and \textit{Fiji}. \textit{3D Slicer} is top project by both rankings (and by
the GithHub stars measure as well).  The Domain Expert ranked \textit{Horos} as
their second choice, while we ranked it twelfth.  Our third ranked project,
\textit{OHIF Viewer} was not listed by the Domain Expert.  Neither were the
software packages that we ranked from fifth to eleventh (\textit{ParaView},
\textit{Weasis}, \textit{medInria}, \textit{BioImage Suite Web}, \textit{OsiriX
Lite}, \textit{INVESALIUS}, and \textit{Gwyddion}).  The software mentioned by
the Domain Expert that we did not rank were the six recommended packages that
did not have visualization as the primary function (as discussed in
Section~\ref{sec_software_selection}).  The differences between the list
recommended by our methodology and the Domain Expert are not surprising.  As
mentioned above, the methodology weights all qualities equally, but that may not
be the case for the Domain Expert's impressions.  Moreover, although the Domain
Expert has significant experience with MI software, he has not used every one of
the 29 packages.

>>>>

To address~\rqref{RQ_CompareHQ2Popular} about how our ranking compares to the
popularity of projects as judged by the scientific community, we make two
comparisons:
\begin{itemize}
\item A comparison of our ranking (from Section~\ref{ch_results}) with the
community ratings on GitHub, as shown by GitHub stars, number of forks, and
number of people watching the projects; and,
\item A comparison of top-rated software from our methodology with the top
recommendations from our domain experts (as mentioned in
Section~\ref{sec_software_selection}).
\end{itemize}

Table~\ref{tab_ranking_vs_GitHub} shows our ranking of the 29 MI projects, and
their GitHub metrics, if applicable. As mentioned in
Section~\ref{sec_score_maintainability}, 24 projects used GitHub. Since GitHub
repositories have different creation dates, we collect the number of months each
stayed on GitHub, and calculate the average number of new stars, people
watching, and forks per 12 months. The method of getting the creation date is
described in Section~\ref{sec_grading_software}.  The items in
Table~\ref{tab_ranking_vs_GitHub} are listed in descending order of the average
number of new stars per year.  The non-GitHub items are listed in the order of
our ranking.  All GitHub statistics were collected in July, 2021.

Generally speaking, most of the top-ranking MI software projects also received
greater attention and popularity on GitHub. Between our ranking and the GitHub
stars-per-year ranking, four of the top five software projects are the same.
Any more stats?

In some cases projects are popular in the community, but our methodology ranked
them low.  This is the case for \textit{dwv}. The reason for the low ranking is
that, as mentioned in Section~\ref{sec_result_installability}, we failed to
build it locally, and used the test version on its websites for the
measurements. We followed the instructions and tried to run the command ``yarn
run test'' locally, which did not work. In addition, the test version did not
detect a broken DICOM file and displayed a blank image as described in
Section~\ref{sec_result_robustness}. We might underestimate the scores for
\textit{dwv} due to uncommon technical issues. We also ranked \textit{DICOM
Viewer} much lower than its popularity. As mentioned in
Section~\ref{sec_result_installability} it depended on the NextCloud platform
that we could not successfully install. Thus, we might underestimate the scores
of its surface reliability and surface robustness. In addition, we weighted all
qualities equally, which is not likely to be the case with all users. As a
result, some projects with high community popularity may have scored lower with
our method because of poor scores for some qualities.  A further explanation for
discrepancies between our measures and the star measures may also be due to
inaccuracy with using stars to approximate popularity.  Stars are not an idea
measure because stars represent the community's feeling in the past more than
current preferences \citep{Szulik2017}, since stars tend only to be added, not
removed.

\begingroup
\renewcommand{\arraystretch}{0.85}
\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
Software & Our ranking & Stars/yr & Watches/yr & Forks/yr \\ 
\midrule
3D Slicer & 1 & 284 & 19 & 128 \\
OHIF Viewer & 3 & 277 & 19 & 224 \\
dwv & 23 & 124 & 12 & 51 \\
ImageJ & 2 & 84 & 9 & 30 \\
ParaView & 5 & 67 & 7 & 28 \\
Horos & 12 & 49 & 9 & 18 \\
Papaya & 17 & 45 & 5 & 20 \\
Fiji & 4 & 44 & 5 & 21 \\
DICOM Viewer & 29 & 43 & 6 & 9 \\
INVESALIUS 3 & 10 & 40 & 4 & 17 \\
Weasis & 6 & 36 & 5 & 19 \\
dicompyler & 26 & 35 & 5 & 14 \\
OsiriX Lite & 9 & 34 & 9 & 24 \\
MRIcroGL & 18 & 24 & 3 & 3 \\
GATE & 25 & 19 & 6 & 26 \\
Ginkgo CADx & 14 & 19 & 4 & 6 \\
BioImage Suite Web & 8 & 18 & 5 & 7 \\
Drishti & 27 & 16 & 4 & 4 \\
Slice:Drop & 20 & 10 & 2 & 5 \\
ITK-SNAP & 15 & 9 & 1 & 4 \\
medInria & 7 & 7 & 3 & 6 \\
SMILI & 13 & 3 & 1 & 2 \\
MatrixUser & 28 & 2 & 0 & 0 \\
MicroView & 16 & 1 & 1 & 1 \\
Gwyddion & 11 & n/a & n/a & n/a \\
XMedCon & 19 & n/a & n/a & n/a \\
DicomBrower & 21 & n/a & n/a & n/a \\
AMIDE & 22 & n/a & n/a & n/a \\
3DimViewer & 24 & n/a & n/a & n/a \\ 
\bottomrule
\end{tabular}
\caption{Software ranking versus GitHub metrics (Sorted in descending order of
GitHub popularity, as estimated by the number of new stars per year)}
\label{tab_ranking_vs_GitHub}
\end{table}
\endgroup

As shown in Section~\ref{sec_software_selection}, our domain experts recommended
a list of top software with 12 software products.  All of the top 4 entries from
the Domain Expert's list are among the top 12 ranked by our methodology. Three
of the top four on both lists are the same: \textit{3D Slicer}, \textit{ImageJ},
and \textit{Fiji}. \textit{3D Slicer} is top project by both rankings (and by
the GithHub stars measure as well).  The Domain Expert ranked \textit{Horos} as
their second choice, while we ranked it twelfth.  Our third ranked project,
\textit{OHIF Viewer} was not listed by the Domain Expert.  Neither were the
software packages that we ranked from fifth to eleventh (\textit{ParaView},
\textit{Weasis}, \textit{medInria}, \textit{BioImage Suite Web}, \textit{OsiriX
Lite}, \textit{INVESALIUS}, and \textit{Gwyddion}).  The software mentioned by
the Domain Expert that we did not rank were the six recommended packages that
did not have visualization as the primary function (as discussed in
Section~\ref{sec_software_selection}).  The differences between the list
recommended by our methodology and the Domain Expert are not surprising.  As
mentioned above, the methodology weights all qualities equally, but that may not
be the case for the Domain Expert's impressions.  Moreover, although the Domain
Expert has significant experience with MI software, he has not used every one of
the 29 packages that were measured.

Although our ranking and the estimate of the community's ranking are not perfect
measures, they do suggest a correlation between best practices and popularity.
We do not know which comes first, the use of best practices or popularity, but
we do know that the top ranked packages tend to incorporate best practices. The
next sections will explore how the practices of the MI community compare to the
broader research software community. We will also investigate the practices from
the top projects that others within the MI community, and within the broader
research software community, can potentially adopt.
